On the discovery of K-enhanced and possibly Mg-depleted stars throughout the Milky Way\footnote{This paper includes data gathered with the 6.5 meter Magellan Telescopes located at Las Campanas Observatory, Chile.}

### 2018-06-09 article/msculled_revised_1.tex
difflatex files from revised_1 and initial submission

% Title of the paper, and the short title which is used in the headers.
% Keep the title short and informative.
\title[Mg--K stars in LAMOST]{On the discovery of K-enhanced and Mg-depleted stars throughout the Milky Way\footnote{This paper includes data gathered with the 6.5 meter Magellan Telescopes located at Las Campanas Observatory, Chile.}}

@@

\begin{abstract}
Stars with unusual elemental abundances offer clues about rare astrophysical events or nucleosynthetic pathways. Stars with significantly depleted magnesium and enhanced potassium ($[{\rm Mg}/{\rm Fe}] < -0.5$; $[{\rm K}/{\rm Fe}] > 1$) have to date only been found in the massive globular cluster NGC~2419 and, to a lesser extent, NGC~2808. The origin of this abundance signature remains unknown, as does the reason for its apparent exclusivity to these two globular clusters. Here we present 112 field stars, identified from \LamostGiants\ \lamost\ giants, that show significantly enhanced [K/Fe] and possibly depleted [Mg/Fe] abundance ratios.
Our sample spans a wide range of metallicities ($-1.5 < [{\rm Fe}/{\rm H}] < 0.3$), yet none show abundance ratios of [K/Fe] or [Mg/Fe] that are as extreme as those observed in NGC~2419. 
If confirmed, the identified sample of stars represents evidence that the nucleosynthetic process producing the anomalous abundances ratios of [K/Fe] and [Mg/Fe] probably occurs at a wide range of metallicities. This would suggest that pollution scenarios that are limited to early epochs (such as Population III supernovae) are an unlikely explanation, although they cannot be ruled out entirely. This sample is expected to help guide modelling attempts to explain the origin of the Mg--K abundance signature, which we now show exists throughout the Milky Way.
\end{abstract}

@@

\section{Method \& Observations}
\label{sec:method}
\subsection{Candidate selection from \lamost\ spectra}
We used a set of \LamostGiants\ giant stars from the second \lamost\ data release \citep{luo2016vizier}. The spectra were placed at rest-frame on a common wavelength sampling from 3905\,\AA\ to 9000\,\AA\ and normalised by \citet{ho2017}. \tc\ \citep{ness2016,ho2017} was used to estimate effective temperature $\teff$, surface gravity $\logg$, metallicity [Fe/H], and mean $\alpha$-element abundance relative to iron [$\alpha$/Fe] using 9,952 stars in common between \lamost\ and the \apogee\ \citep{alam2015} survey. Unless otherwise stated, these transferred labels are those referred to throughout this study. For details regarding the preparatory work, model generation, and label transfer between the \apogee\ and \lamost\ surveys we direct the reader to \citet{ho2017}.

@@

We fit a Gaussian profile to the flux residuals for all three absorption lines in the Mg triplet (5167\,\AA, 5172\,\AA, 5184\,\AA) as well as the K doublet (7665\,\AA, 7699\,\AA) for all \LamostGiants\ giants. For each star we recorded the amplitudes, widths (standard deviations) and wavelength (mean) about which the Gaussian profiles were fit. The amplitude was used as the measure of the discrepancy from the model spectra, and is analogous to the depth of the absorption line, and in principle linearly dependent on the equivalent width. This measure was favoured as more reliable than attempting to quantify line widths from the low resolution \lamost\ spectra. %The relationship between equivalent width ($EW$), the amplitude ($A$) and the Gaussian width ($\sigma$) is given by $EW=\frac{2\pi A}{\sigma}$.

We identified candidates by requiring that they match at least one of the following three quality filters:
\begin{enumerate}
\item We required the amplitude $A$ of the profile at the Mg 5184 \AA \ line to satisfy $A_{{\rm Mg} @ 5184} > 0.05$ and the amplitude of the profile at the K 7665\,\AA\ line to satisfy $A_{{\rm K}\,@\,7665} < -0.05$. Both amplitudes must also be measured at more than $3\sigma$ ($|A|/\sigma_{A} \geq 3$).
\item We required any two of the three Mg triplet lines to satisfy $A > 0.05$ and at least one K line to have $A < -0.05$, and for those amplitudes to have $|A|/\sigma_{A} \geq 3$.
\item We required any two of the three Mg triplet lines to have $A > 0$ and both K lines to have $A < 0$, and for the spectra to have a signal-to-noise ($S/N$) ratio of $S/N > 30$ in \lamost\ and a reported $\chi_{r}^2 < 3$ from \tc\ \citep{ness2016,ho2017}.
\end{enumerate}

@@

There are subtleties to this analysis which likely have no significant bias in our inferred abundances, but we include here for completeness. Specifically, we assume that some stellar absorption is already accounted for by our data-driven model. In a sense, the data-driven model predicts a `typical' spectrum for a star with the given stellar parameters. This `typical` spectrum will include some contribution from all abundances (e.g., Mg, K, Al, Na), which is encapsulated by the overall metallicity [Fe/H]. That is to say that for a star of [Fe/H]$ = 0$, our data-driven model is predicting the spectrum of a \emph{typical} star at $[{\rm Fe/H}] = 0$, and that typical star would have $[{\rm K/Fe}] \sim 0$. We synthesise spectra to account for the flux residuals \emph{away} from the data-driven model. 
%@andy
% vt = 2.700 ? 0.509 log g km/s. (7

@@

Our sample of \lamost\ giants carries with it certain selection effects from both the \apogee\ DR12 and \lamost\ samples. The scope of this the data-driven model is necesarily confined to the training sample, which is in turn confined in this case to stars common between \lamost\ and \apogee\. For example, metal-poor stars with [Fe/H] $\lesssim-2$ are mostly absent from our sample because they were too dissimilar from the training stars for label transfer from \apogee. Similarly, the target selection process for \lamost\ is not easily invertible so we do not consider selection effects present in \lamost\ beyond comparing the candidate stars' metallicity distribution with that of \cite{ho2017}. We find that the frequency of Mg--K candidate stars is consistent (within the errors) of being flat with metallicity.

@@

The modelling work of \cite{prantzos2017} predicted an anti-correlation between Na and K in a 180\,MK hydrogen burning environment, while observational data for NGC 2808 implies a positive correlation between Na and K. Upon examining the [Na/Fe] abundance ratios we estimated from \lamost\ spectra of our candidates, based on the doublet around 5850 \AA, we observed two distinct populations: one population making up approximately 30 percent of the sample with [Na/Fe] $> 0.5$, significantly higher than typical Galactic abundances of [Na/Fe] $\lesssim$ 0.3 \citep{kobayashi2011}, and a second population with [Na/Fe] abundance ratios consistent with typical Galactic levels. The two groups can be clearly distinguished in Figure \ref{NavsFeh}, where for comparison we also show the abundances of [Na/Fe] versus [Fe/H] from the \apogee\ survey.

@@

Given that the sample is almost certainly comprised of disk stars, it is relevant to also compare with typical Galactic abundances for [K/Fe]. In Figure \ref{KvsMg} and \ref{KvsFe} we make these comparisons with data from the \apogee\ survey \citep{alam2015}. The sample's [K/Fe] appears definitively anomalous, with Galactic [K/Fe] around $0.3 \pm 0.2$\,dex, while the identified candidate stars have a spread in [K/Fe] from around 0.8 to 1.3 dex.

@@

The implications of the abundance ratios we find are intriguing. The fact that in such a large sample of \LamostGiants\ field stars, none were found that replicated (or exceeded) the extreme abundance pattern of NGC 2419. This highlights the uniqueness of NGC 2419.

However, we repeat that our sample of \lamost\ giants searched had already been filtered of most of the stars that would have matched NGC 2419's metallicity. The 112 star candidate pool identified of K enhanced stars represents a unique collection of stars; while not as K enhanced as the extreme population in NGC 2419, they exhibit K abundances well above typical Galactic levels at a range of metallicities. It seems likely that whatever process is responsible for these anomalous field stars is similar to whatever caused the unusual stellar populations in NGC 2419 and NGC 2808.  Although a unifying explanation remains elusive, it is improbable that a processes acting to produce such high levels of [K/Fe] in the Milky Way is unrelated to whatever process is responsible for the extreme abundance ratios in NGC 2419. The apparent consistency of NGC 2419's extreme population with the anti-correlation between [K/Fe] and [Fe/H] observed in Figure \ref{KvsFe} further supports the idea that the process responsible for the stars in this study and the previously known anomalous stars in NGC 2419 are related.

@@

Our study presents a collection of stars in the disk of the Galaxy over a range of metallicities that show anomalously high levels of [K/Fe]. The spread in metallicities among our candidates is important. It implies that the process responsible for the Mg--K abundance signature may not be tied to a certain epoch in the Milky Way's evolution. These stars with enhanced potassium apparently tend to form at all ages, across all metallicities, so it is probable that the polluting source is also independent of metallicity. If we assume that the polluter star forms at a range of metallicities, then we can attempt to constrain scenarios that may be responsible for the Mg--K anti-correlation by disregarding polluters that are reliant on specific environments to form (e.g., low metallicity environments).

One class of candidates that can be immediately eliminated under such an assumption are Population III stars. A hyper-massive Population III star may be been a plausible explanation for NGC~2419 \citep{carretta2013} exclusively, particularly given its isolation from the Milky Way ($d_{\rm GC} \approx 90\,{\rm kpc}$) and low metallicity. However, being unable to form throughout the life of the Galaxy makes it unlikely for Population III stars to be the polluters responsible for the stars we identify.

>>>>

% Title of the paper, and the short title which is used in the headers.
% Keep the title short and informative.
\title[Mg--K stars in LAMOST]{On the discovery of K-enhanced and possibly Mg-depleted stars throughout the Milky Way\footnote{This paper includes data gathered with the 6.5 meter Magellan Telescopes located at Las Campanas Observatory, Chile.}}

@@

\begin{abstract}
Stars with unusual elemental abundances offer clues about rare astrophysical events or nucleosynthetic pathways. Stars with significantly depleted magnesium and enhanced potassium ($[{\rm Mg}/{\rm Fe}] < -0.5$; $[{\rm K}/{\rm Fe}] > 1$) have to date only been found in the massive globular cluster NGC~2419 and, to a lesser extent, NGC~2808. The origin of this abundance signature remains unknown, as does the reason for its apparent exclusivity to these two globular clusters. Here we present 112 field stars, identified from \LamostGiants\ \lamost\ giants, that show significantly enhanced [K/Fe] and possibly depleted [Mg/Fe] abundance ratios.
Our sample spans a wide range of metallicities ($-1.5 < [{\rm Fe}/{\rm H}] < 0.3$), yet none show abundance ratios of [K/Fe] or [Mg/Fe] that are as extreme as those observed in NGC~2419. 
If confirmed, the identified sample of stars represents evidence that the nucleosynthetic process producing the anomalous abundances ratios of [K/Fe] and [Mg/Fe] probably occurs at a wide range of metallicities. This would suggest that pollution scenarios that are limited to early epochs (such as Population III supernovae) are an unlikely explanation, although they cannot be ruled out entirely. This sample is expected to help guide modelling attempts to explain the origin of the Mg--K abundance signature.
\end{abstract}

@@

\section{Method \& Observations}
\label{sec:method}
\subsection{Candidate selection from \lamost\ spectra}
We used a set of \LamostGiants\ giant stars from the second \lamost\ data release \citep{luo2016vizier}. The spectra were placed at rest-frame on a common wavelength sampling from 3905\,\AA\ to 9000\,\AA\ and normalised by \citet{ho2017}. \tc\ \citep{ness2016,ho2017} was used to estimate effective temperature $\teff$, surface gravity $\logg$, metallicity [Fe/H], and mean $\alpha$-element abundance relative to iron [$\alpha$/Fe] using 9,952 stars in common between \lamost\ and the \apogee\ \citep{alam2015} survey. This process is referred to as 'label transfer'. Unless otherwise stated, these transferred labels are those referred to throughout this study. For details regarding the preparatory work, model generation, and label transfer between the \apogee\ and \lamost\ surveys we direct the reader to \citet{ho2017}.

@@

We fit a Gaussian profile of amplitude $A$ to the flux residuals for all three absorption lines in the Mg triplet (5167\,\AA, 5172\,\AA, 5184\,\AA) as well as the K doublet (7665\,\AA, 7699\,\AA) for all \LamostGiants\ giants. For each star we recorded the amplitudes, widths (standard deviations) and wavelength (mean) about which the Gaussian profiles were fit. The amplitude was used as the measure of the discrepancy from the model spectra, and is analogous to the depth of the absorption line, and in principle linearly dependent on the equivalent width. This measure was favoured as more reliable than attempting to quantify line widths from the low resolution \lamost\ spectra. %The relationship between equivalent width ($EW$), the amplitude ($A$) and the Gaussian width ($\sigma$) is given by $EW=\frac{2\pi A}{\sigma}$.

We identified candidates by requiring that they match at least one of the following three quality filters:
\begin{enumerate}
\item We required the amplitude $A$ of the profile at the Mg 5184 \AA \ line to satisfy $A_{{\rm Mg} @ 5184} > 0.05$ and the amplitude of the profile at the K 7665\,\AA\ line to satisfy $A_{{\rm K}\,@\,7665} < -0.05$. Both amplitudes must also be measured at more than $3\sigma$ ($|A|/\sigma_{A} \geq 3$).
\item We required any two of the three Mg triplet lines to satisfy $A > 0.05$ and at least one K line to have $A < -0.05$, and for those amplitudes to have $|A|/\sigma_{A} \geq 3$.
\item We required the amplitude of any two of the three Mg triplet lines to have $A > 0$ and both K lines to have $A < 0$, and for the spectra to have a signal-to-noise ($S/N$) ratio of $S/N > 30$ in \lamost\ and a reported $\chi_{r}^2 < 3$ from \tc\ \citep{ness2016,ho2017}.
\end{enumerate}

@@

It can be assumed that some stellar absorption is already accounted for by our data-driven model. In a sense, the data-driven model predicts a `typical' spectrum for a star with the given stellar parameters. This `typical` spectrum will include some contribution from all abundances (e.g., Mg, K, Al, Na), which is encapsulated by the overall metallicity [Fe/H]. That is to say that for a star of [Fe/H]$ = 0$, our data-driven model is predicting the spectrum of a \emph{typical} star at $[{\rm Fe/H}] = 0$, and that typical star would have $[{\rm K/Fe}] \sim 0$. We synthesise spectra to account for the flux residuals \emph{away} from the data-driven model. 
%@andy
% vt = 2.700 ? 0.509 log g km/s. (7

@@

Our sample of \lamost\ giants carries with it certain selection effects from both the \apogee\ DR12 and \lamost\ samples. The scope of this the data-driven model is necesarily confined to the training sample, which is in turn confined in this case to stars common between \lamost\ and \apogee.\ For example, metal-poor stars with [Fe/H] $\lesssim-2$ are mostly absent from our sample because they were too dissimilar from the training stars for label transfer from \apogee. Similarly, the target selection process for \lamost\ is not easily invertible so we do not consider selection effects present in \lamost\ beyond comparing the candidate stars' metallicity distribution with that of \cite{ho2017}. We find that the frequency of Mg--K candidate stars is consistent (within the errors) of being flat with metallicity.

@@

The modelling work of \cite{prantzos2017} predicted an anti-correlation between Na and K which could result from a 180\,MK hydrogen burning environment, while observational data for NGC 2808 implies a positive correlation between Na and K. Upon examining the [Na/Fe] abundance ratios we estimated from \lamost\ spectra of our candidates, based on the doublet around 5850 \AA, we observed two distinct populations: one population making up approximately 30 percent of the sample with [Na/Fe] $> 0.5$, significantly higher than typical Galactic abundances of [Na/Fe] $\lesssim$ 0.3 \citep{kobayashi2011}, and a second population with [Na/Fe] abundance ratios consistent with typical Galactic levels. The two groups can be clearly distinguished in Figure \ref{NavsFeh}, where for comparison we also show the abundances of [Na/Fe] versus [Fe/H] from the \apogee\ survey.

@@

Given that the sample is almost certainly comprised of disk stars, it is relevant to also compare with typical Galactic abundances for [K/Fe]. In Figure \ref{KvsMg} and \ref{KvsFe} we make these comparisons with data from the \apogee\ survey \citep{alam2015}. Our sample's [K/Fe] appears definitively anomalous, with Galactic [K/Fe] around $0.3 \pm 0.2$\,dex, while the identified candidate stars have a spread in [K/Fe] from around 0.8 to 1.3 dex.

@@

The implications of the abundance ratios we find are intriguing. It is an astonishing fact that in such a large sample of \LamostGiants\ field stars, none were found that replicated (or exceeded) the extreme abundance pattern of NGC 2419 highlights the uniqueness of NGC 2419.

However, we repeat that the sample of \lamost\ giants  we searched had already been filtered of most of the stars that would have matched NGC 2419's metallicity. The 112 star candidate pool identified of K enhanced stars represents a unique collection of stars; while not as K enhanced as the extreme population in NGC 2419, they exhibit K abundances well above typical Galactic levels at a range of metallicities. It seems likely that whatever process is responsible for these anomalous field stars is similar to whatever caused the unusual stellar populations in NGC 2419 and NGC 2808.  Although a unifying explanation remains elusive, it is improbable that a processes acting to produce such high levels of [K/Fe] in the Milky Way is unrelated to whatever process is responsible for the extreme abundance ratios in NGC 2419. The apparent consistency of NGC 2419's extreme population with the anti-correlation between [K/Fe] and [Fe/H] observed in Figure \ref{KvsFe} further supports the idea that the process responsible for the stars in this study and the previously known anomalous stars in NGC 2419 are related.

@@

Our study presents a collection of stars in the disk of the Galaxy with a range of metallicities that show anomalously high levels of [K/Fe]. The spread in metallicities among our candidates is important. It implies that the process responsible for the Mg--K abundance signature may not be tied to a certain epoch in the Milky Way's evolution. These stars with enhanced potassium apparently tend to form at all ages, across all metallicities, so it is probable that the polluting source is extant at all metallicities. If we assume that the polluter star forms at a range of metallicities, then we can attempt to constrain scenarios that may be responsible for the Mg--K anti-correlation by disregarding polluters that are reliant on specific environments to form (e.g., low metallicity environments).

One class of candidates that can be immediately eliminated under such an assumption are Population III stars. A hyper-massive Population III star may be been a plausible explanation for NGC~2419 \citep{carretta2013} exclusively, particularly given its isolation from the Milky Way at a distance of $\approx 90\,{\rm kpc}$ and low metallicity. However, being unable to form throughout the life of the Galaxy makes it unlikely for Population III stars to be the polluters responsible for the stars we identify.
